### 2021-09-02 input/abstract.tex
typos of review

\begin{abstract}
  %% Context and problem
  Although the holy grail to store and manipulate data in Edge
  infrastructures is yet to be found, state-of-the-art approaches
  demonstrated the relevance of replication strategies that bring
  contents closer to consumers: The latter enjoy better response time
  while the volume of data passing through the network decreases
  overall.
  %% Although, there is no major solution to store and manipulate data in
  %% Edge infrastructures yet, existing proposals demonstrated the
  %% relevance of appropriate replication strategies to bring contents
  %% closer to consumers, reducing the impact of the latency as well as
  %% the volume of data passing through the network overall.  
  Unfortunately, locating the closest replica of a specific content
  requires indexing \emph{every live} replica along with its
  \emph{location}.
  %
  %Current proposals built on a centralized service face scalability
  %issues, single point of failure, as well as latency penalities where
  %locating replicas may effectively take more time than actually
  %downloading contents. Instead, data consumer \processes must
  %maintain an up-to-date index themselves. Unfortunately, maintaining
  %an index of all live replicas at every \process would prove overly
  %costly in terms of memory and generated traffic, especially in large
  %scale dynamic systems where \processes create and destroy replicas
  %at any time.
  % 
  Relying on remote services contradicts with the properties of Edge
  infrastructures as locating replicas may effectively take more time
  than actually downloading contents.  At the opposite, maintaining
  such an index at every \process would prove overly costly in terms
  of memory and traffic, especially since \processes can create and
  destroy replicas at any time.

>>>>

\begin{abstract}
  %% Context and problem
  Although the holy grail to store and manipulate data in Edge
  infrastructures is yet to be found, state-of-the-art approaches
  demonstrated the relevance of replication strategies that bring
  content closer to consumers: The latter enjoy better response time
  while the volume of data passing through the network decreases
  overall.
  %% Although, there is no major solution to store and manipulate data in
  %% Edge infrastructures yet, existing proposals demonstrated the
  %% relevance of appropriate replication strategies to bring contents
  %% closer to consumers, reducing the impact of the latency as well as
  %% the volume of data passing through the network overall.  
  Unfortunately, locating the closest replica of a specific content
  requires indexing \emph{every live} replica along with its
  \emph{location}.
  %
  %Current proposals built on a centralized service face scalability
  %issues, single point of failure, as well as latency penalities where
  %locating replicas may effectively take more time than actually
  %downloading contents. Instead, data consumer \processes must
  %maintain an up-to-date index themselves. Unfortunately, maintaining
  %an index of all live replicas at every \process would prove overly
  %costly in terms of memory and generated traffic, especially in large
  %scale dynamic systems where \processes create and destroy replicas
  %at any time.
  % 
  Relying on remote services enters in contradiction with the
  properties of Edge infrastructures as locating replicas may
  effectively take more time than actually downloading content.  At
  the opposite, maintaining such an index at every \process would
  prove overly costly in terms of memory and traffic, especially since
  \processes can create and destroy replicas at any time.

### 2021-09-02 input/background.tex
intro reworkd with bullet and contribution
cosmetic + number
first step in styling
typos of review

@@

Then, in Figure~\ref{fig:partition_intuitionB}, another \Process~$G$
creates a replica of this content. Similarly to \Process~$R$, it
notifies other \processes of this new replica. However, to avoid that
north \processes request its replica, and south \processes request the
northern one, \processes hosting a replica must add \emph{location
information} along with their notifications. As consequence, every
\process eventually knows every replica location and can download from
its closest host. \underline{R}ed \processes would request
\Process~$R$ while \underline{g}reen \processes would request
\Process~$G$.

@@

Then, in Figure~\ref{fig:partition_intuitionC}, another \Process~$B$
creates a replica of this content. Similarly to \Process~$R$ and
\Process~$G$, \Process~$B$ can notify other \processes of this new
replica. However, the set of \processes that could actually use this
new replica is actually much smaller than the network size. Uniform
reliable broadcast is not designed for such a context and would
generate a lot of unnecessary traffic. Instead, \processes need a
communication primitive that propagates notifications within a scope
by evaluating a predicate, starting at its broadcaster (the
\emph{source}).  In other terms, \processes propagate notifications as
long as they consider them useful based on location information they
carry. We call such primitive \emph{scoped broadcast} (see
Section~\ref{subsec:scoped}), for messages transitively reach a subset
of interconnected \processes (the \emph{scope}). Using this primitive,
\processes can lock down the traffic of content indexing to relevant
\processes.

@@

Finally, in Figure~\ref{fig:partition_intuitionD}, \Process~$G$
destroys its replica. Every \processes that belonged to its green
partition must choose another partition to be in. While it makes sense
for \Process~$G$ to scoped broadcast its removal, \Process~$B$ and
\Process~$R$ cannot afford to continuously advertise their replica to
fill the gap left open by \Process~$G$. Instead, we want to trigger
once again scoped broadcast at bordering \processes of red and blue
partitions. The scope of scoped broadcast changes over receipts by
\processes. Dynamic partitioning raises additional challenges related
to concurrent operations where removed partitions would block the
propagation of other partitions (see Section~\ref{subsec:dynamic}).
\end{asparadesc}

@@

\vfill
\TODO{spacing.}
\newpage

>>>>

\input{input/figpartitionintuition.tex} %% positioning

@@

Then, in Figure~\ref{fig:partition_intuitionB}, another \Process~$G$
creates a replica of this content. Similarly to \Process~$R$, it
notifies other \processes of this new replica. However, to avoid that
north \processes request its replica, and south \processes request the
northern one, \processes hosting a replica must add \emph{location
information} along with their notifications. As a consequence, every
\process eventually knows every replica location and can download from
its closest host. \underline{R}ed \processes would request
\Process~$R$ while \underline{g}reen \processes would request
\Process~$G$.

@@

Then, in Figure~\ref{fig:partition_intuitionC}, another \Process~$B$
creates a replica of this content. Similarly to \Process~$R$ and
\Process~$G$, \Process~$B$ can notify other \processes of this new
replica. However, the set of \processes that could actually use this
new replica is actually much smaller than the network size. Uniform
reliable broadcast is not designed for such a context and would
generate a lot of unnecessary traffic. Instead, \processes need a
communication primitive that propagates notifications within a scope
by evaluating a predicate, starting at its broadcaster (the
\emph{source}).  In other terms, \processes propagate notifications as
long as they consider them useful based on location information they
carry. We call such a primitive \emph{scoped broadcast} (see
Section~\ref{subsec:scoped}), for messages transitively reach a subset
of interconnected \processes (the \emph{scope}). Using this primitive,
\processes can lock down the traffic of content indexing to relevant
\processes.

@@

Finally, in Figure~\ref{fig:partition_intuitionD}, \Process~$G$
destroys its replica. Every \process that belonged to its green
partition must choose another partition to be in. While it makes sense
for \Process~$G$ to scoped broadcast its removal, \Process~$B$ and
\Process~$R$ cannot afford to continuously advertise their replica to
fill the gap left open by \Process~$G$. Instead, we want to trigger
once again scoped broadcast at bordering \processes of red and blue
partitions. The scope of scoped broadcast changes over receipts by
\processes. Dynamic partitioning raises additional challenges related
to concurrent operations where removed partitions would block the
propagation of other partitions (see Section~\ref{subsec:dynamic}).
\end{asparadesc}

@@

### 2021-09-02 input/conclusion.tex
typos of review

With the advent of the Edge and IoT era, major distributed services
proposed by our community should be revised in order to mitigate as
much as possible traffic between \processes.  In this paper, we
tackled the problem of content indexing, a key service for storage
systems.
%
We proposed to address this problem as a dynamic logical partitioning
where partitions grow and shrink to reflect content and replicas
locations, as well as infrastructure changes. Using an efficient
communication primitive called scoped broadcast, each \process
composing the infrastructure eventually knows the \process hosting the
best replica of specific content.  The challenge resides in handling
concurrent operations that may impair the propagation of messages, and
in turns, lead to inconsistent partitioning.
%% The major difficulty lies in concurrent operations, each having a
%% different scope, and ensure the correct convergence of the system.  We
%% defined, in particular, multiple properties that highlight this
%% complexity.
%
We highlighted the properties that solve this problem and proposed an
implementation called \NAME.  Our simulations confirmed that
\processes quickly reach consistent partitioning together, and that
the generated traffic remains locked down to partitions.
%% correctness of our protocol as well as the confinement of
%% the traffic only where it is needed.

@@

\noindent We also would like to evaluate \NAME within a concrete
storage systems such as \underline{I}nter\underline{P}lanetary
\underline{F}ile \underline{S}ystem
(IPFS)~\cite{henningsen2020mapping}. This would allow us to assess the
relevance of \NAME in a real system subject to high dynamicity, and
compare it against its current DHT-based indexing system that does not
include distance in its operation.

>>>>

With the advent of the Edge and IoT era, major distributed services
proposed by our community should be revised in order to mitigate as
much as possible traffic between \processes.  In this paper, we
tackled the problem of content indexing, a key service for storage
systems.
%
We proposed to address this problem as a dynamic logical partitioning
where partitions grow and shrink to reflect content and replicas
locations, as well as infrastructure changes. Using an efficient
communication primitive called scoped broadcast, each \process
composing the infrastructure eventually knows the \process hosting the
best replica of specific content.  The challenge resides in handling
concurrent operations that may impair the propagation of messages, and
in turn, lead to inconsistent partitioning.
%% The major difficulty lies in concurrent operations, each having a
%% different scope, and ensure the correct convergence of the system.  We
%% defined, in particular, multiple properties that highlight this
%% complexity.
%
We highlighted the properties that solve this problem and proposed an
implementation called \NAME.  Our simulations confirmed that
\processes quickly reach consistent partitioning together, and that
the generated traffic remains locked down to partitions.
%% correctness of our protocol as well as the confinement of
%% the traffic only where it is needed.

@@

\noindent We also would like to evaluate \NAME within a concrete
storage system such as \underline{I}nter\underline{P}lanetary
\underline{F}ile \underline{S}ystem
(IPFS)~\cite{henningsen2020mapping}. This would allow us to assess the
relevance of \NAME in a real system subject to high dynamicity, and
compare it against its current DHT-based indexing system that does not
include distance in its operation.

### 2021-09-02 input/experiments.tex
typos of review

This section describes the results of our experimentation evaluating
the properties of \NAME that enables adaptive scoped broadcast in
dynamic networks. It answers the following questions: What is the
overhead of reaching consistent partitioning using \NAME in large
scale networks?  Does \NAME is relevant in Edge infrastructures where
\processes are geo-distributed into clusters? We performed the
experimentation using PeerSim, a simulator written in Java that allows
researchers to evaluate and assess distributed algorithms in large
scale networks~\cite{montresor2009peersim}. The code is available on
the Github platform at
\url{https://anonymous.4open.science/r/peersim-partition-5592}.

@@

Figure~\ref{fig:geant} shows the results of this experimentation. The
top part displays the one source experiment while the bottom part
displays the one source per cluster experiment.

>>>>

This section describes the results of our experimentation evaluating
the properties of \NAME that enables adaptive scoped broadcast in
dynamic networks. It answers the following questions: What is the
overhead of reaching consistent partitioning using \NAME in large
scale networks?  Is \NAME relevant in Edge infrastructures where
\processes are geo-distributed into clusters? We performed the
experimentation using PeerSim, a simulator written in Java that allows
researchers to evaluate and assess distributed algorithms in large
scale networks~\cite{montresor2009peersim}. The code is available on
the Github platform at
\url{https://anonymous.4open.science/r/peersim-partition-5592}.

@@

Figure~\ref{fig:geant} shows the results of this experimentation. The
top part displays the results with one source while the bottom part
displays the results with one source per cluster.

### 2021-09-02 input/introduction.tex
cosmetic + number
first step in styling
typos of review

\input{input/figpartitionintuition.tex} %% positioning

@@

These approaches contradict with the objectives of Edge
infrastructures that aim at reducing the impact of latency as well as
the volume of data passing through the network.
%
First, accessing a remote \node to request content location(s) raises
hot spots and availability issues. But most importantly, it results in
additional delays~\cite{asrese2019measuring, doan2019tracing} that
occur even before the actual download started.
%
Second, the client gets a list of content locations at the discretion
of content indexing services. Without information about these
locations, it often ends up downloading from multiple replica hosts,
yet only keeping the fastest answer. In turns, clients either waste
network resources, or face lower response time.
%% there is no information to select the ``best'' node among the
%% possible candidates from which the content could be retrieved. In
%% most cases, this results in parallel requests performed by the
%% client node to each candidate with the goal of retrieving the
%% content as efficient as possible.

To tackle aforementioned limitations, every \process that might
request or replicate content must also host its own content indexing
service in a fully decentralized fashion~\cite{kermarrec2015want}. At
any time, it can immediately locate the closest replica of specific
content.  A naive approach would be that every \process indexes and
ranks \emph{every live} replica along with its \emph{location}
information. When creating or destroying a replica, a \process would
notify all other \processes by efficiently broadcasting its
operation~\cite{birman1999bimodal, hadzilacos1994modular,
  raynal2013distributed}. Unfortunately, this also contradicts with
Edge infrastructure objectives, for such protocol does not confine the
traffic generated to maintain its indexes. A \process may acknowledge
the existence of replicas at the other side of the network while there
already exists a replica next to it.

@@

The rest of this paper is organized as
follows. Section~\ref{sec:background} illustrates the motivation and
problem behind our proposal.  Section~\ref{sec:adaptive} describes
dynamic consistent partitioning and its implementation along with its
complexity. Section~\ref{sec:experimentation} presents our
evaluations.  Section~\ref{sec:related_work} reviews related
work. Finally, Section~\ref{sec:conclusion} concludes and discusses
future work.

>>>>

@@

These approaches are in contradiction with the objectives of Edge
infrastructures that aim at reducing the impact of latency as well as
the volume of data passing through the network.
%
First, accessing a remote \node to request content location(s) raises
hot spots and availability issues. But most importantly, it results in
additional delays~\cite{asrese2019measuring, doan2019tracing} that
occur even before the actual download started.
%
Second, the client gets a list of content locations at the discretion
of content indexing services. Without information about these
locations, it often ends up downloading from multiple replica hosts,
yet only keeping the fastest answer. In turn, clients either waste
network resources, or face lower response time.
%% there is no information to select the ``best'' node among the
%% possible candidates from which the content could be retrieved. In
%% most cases, this results in parallel requests performed by the
%% client node to each candidate with the goal of retrieving the
%% content as efficient as possible.

To address the aforementioned limitations, every \process that might
request or replicate content must also host its own content indexing
service in a fully decentralized fashion~\cite{kermarrec2015want}. At
any time, it can immediately locate the closest replica of specific
content.  A naive approach would be that every \process indexes and
ranks \emph{every live} replica along with its \emph{location}
information. When creating or destroying a replica, a \process would
notify all other \processes by efficiently broadcasting its
operation~\cite{birman1999bimodal, hadzilacos1994modular,
  raynal2013distributed}. Unfortunately, this also contradicts Edge
infrastructure objectives, for such a protocol does not confine the
traffic generated to maintain its indexes. A \process may acknowledge
the existence of replicas at the other side of the network while there
already exists a replica next to it.

@@

The rest of this paper is organized as
follows. Section~\ref{sec:background} illustrates the motivation and
problem behind our proposal.  Section~\ref{sec:adaptive} describes
dynamic consistent partitioning and its implementation along with an
analysis of its complexity. Section~\ref{sec:experimentation} presents
our evaluations.  Section~\ref{sec:related_work} reviews related
work. Finally, Section~\ref{sec:conclusion} concludes and discusses
future work.

### 2021-09-02 input/proposal.tex
typos of review

To assign and maintain \processes to their best partition according to
replica creations and removals, as well as dynamic infrastructure
changes, we designed and implemented \NAME.  \NAME stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}. It
relies on a primitive that allows a \process to broadcast a message
within a limited scope. We first use this primitive to guarantee
consistent partitioning when a \process can only create a new replica
within the system. We highlight the issue when a process can also
destroy a replica, and provide a second algorithm that handles replica
removals as well as dynamic changes of the infrastructure.  This
section describes the communication primitive called scoped broadcast,
then discusses the properties that guarantee consistent partitioning
in our context, and finally details our implementation \NAME and its
complexity.

@@

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency.

@@

\begin{definition}[\label{def:lwo}Last win order and staleness]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m$
  convey \emph{stale} control information.
\end{definition}

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such a solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belong to in the
past. This would prove overly expensive in dynamic systems where
\processes join and leave the system, create and delete partitions, at
any time.  Instead, we propose to use the delivery order at each
\process to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, f, g]$ delivered
  and forwarded a message $\alpha_d$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from \process to \process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_h^z$ with last
    $\alpha_h^z$, for $\alpha_h^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ that do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_h$ if need
    be. Nonetheless, they deliver and forward $\alpha_h$ that at least
    one following \process will receive:
  \item [\processes with $\delta_h \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ that do not belong to the preceding category,
    for they already delivered $\delta_h$. Nevertheless, best eventual
    forwarding guarantees that they receive $\alpha_h$, which
    contradicts their history or state. In such a case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted. Indeed their parent may have
    blocked the corresponding $\delta$ messages and subsequent
    \processes must purge their possibly stale state, for the sake of
    consistency.  We note such notification $\delta_d?$ to emphasize
    the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$g$ that have
    $\alpha_d$ but preceding delivered messages are not
    important. Such \processes must trust detecting \processes by
    delivering and forwarding $\delta_d?$. It is worth noting that
    these messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source. Each message $m$ carries its source $s$ along with its version
$c$ that we subscript $m_{s, c}$.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  Its size
increases linearly with the number of sources that the \process ever
known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta_{s, c}^{\pi}$ message implements the uncertain $\delta?$
message of Figure~\ref{fig:proofB}. It fulfils the double role of
removing stale control information without modifying the local counter
of the source (since the partition may still exist), and removing
loops to ensure termination (since \processes do not modify their
local state to reflect such a delivery, messages piggyback
monotonically such increasing control information).

@@

\NAME guarantees dynamic consistent partitioning by making extensive
use of \process-to-\process communications. It requires
\begin{inparaenum}[(i)]
\item purging stale control information, hence propagating $\delta$
  messages; and
\item retrieving the closest source, hence propagating $\alpha$
  messages.
\end{inparaenum}
In terms of number of messages, in the average-case, a \process $i$
chosen uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half in average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process. % \TODO{Multiple receipt and
multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.

>>>>

To assign and maintain \processes to their best partition according to
replica creations and removals, as well as dynamic infrastructure
changes, we designed and implemented \NAME.  \NAME stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}. It
relies on a primitive that allows a \process to broadcast a message
within a limited scope. We first use this primitive to guarantee
consistent partitioning when a \process can only create a new replica
within the system. We highlight the issue when a process can also
destroy a replica, and provide a second algorithm that handles replica
removals as well as dynamic changes of the infrastructure.  This
section describes the communication primitive called scoped broadcast,
then discusses the properties that guarantee consistent partitioning
in our context, and finally details our implementation \NAME and the
analysis of its complexity.

@@

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an exttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such a distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, this would be about link latency.

@@

\begin{definition}[\label{def:lwo}Last win order and staleness]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\nexists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m$
  convey \emph{stale} control information.
\end{definition}

@@

One could guarantee consistent partitioning by always propagating the
$\delta$ messages corresponding to the $\alpha$ messages it propagated
before. In Figure~\ref{fig:problem}, it means that as soon as
\Process~$b$ forwards $\alpha_a^2$, it assumes that its neighbors
\Process~$c$ and \Process~$d$ may need the notification of removal
$\delta_a$ if it exists. However, such a solution also implies that
\processes generate traffic not only related to their current
partition, but also related to partitions they belonged to in the
past. This would prove overly expensive in dynamic systems where
\processes join and leave the system, create and delete partitions, at
any time.  Instead, we propose to use the delivery order at each
\process to detect possible inconsistencies and solve them. Together,
\processes eventually remove all stale control information (transiting
messages and local states) of the system leaving room for propagation
of messages about up-to-date partitions.

@@

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, f, g]$ delivered
  and forwarded a message $\alpha_d$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from \process to \process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_h^z$ with last
    $\alpha_h^z$, for $\alpha_h^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ that do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_h$ if need
    be. Nonetheless, they deliver and forward $\alpha_h$ that at least
    one following \process will receive.
  \item [\processes with $\delta_h \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ that do not belong to the preceding category,
    for they already delivered $\delta_h$. Nevertheless, best eventual
    forwarding guarantees that they receive $\alpha_h$, which
    contradicts their history or state. In such a case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted. Indeed their parent may have
    blocked the corresponding $\delta$ messages and subsequent
    \processes must purge their possibly stale state, for the sake of
    consistency.  We note such a notification $\delta_d?$ to emphasize
    the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$g$ that have
    $\alpha_d$ but preceding delivered messages are not
    important. Such \processes must trust detecting \processes by
    delivering and forwarding $\delta_d?$. It is worth noting that
    these messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source. Each message $m$ carries its source $s$ along with its version
$c$ that we subscript $m_{s, c}$.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  Its size
increases linearly with the number of sources that the \process has
ever known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such a
$\delta_{s, c}^{\pi}$ message implements the uncertain $\delta?$
message of Figure~\ref{fig:proofB}. It fulfils the double role of
removing stale control information without modifying the local counter
of the source (since the partition may still exist), and removing
loops to ensure termination (since \processes do not modify their
local state to reflect such a delivery, messages piggyback
monotonically such increasing control information).

@@

\NAME guarantees dynamic consistent partitioning by making extensive
use of \process-to-\process communications. It requires
\begin{inparaenum}[(i)]
\item purging stale control information, hence propagating $\delta$
  messages; and
\item retrieving the closest source, hence propagating $\alpha$
  messages.
\end{inparaenum}
In terms of number of messages, in the average case, a \process $i$
chosen uniformly at random among all \processes creates a logical
partition. Its messages $\alpha_i$ spread through the network until
reaching \processes that belong to another partition closer to
them. This splits partitions in half on average. Overall, the $a^{th}$
new partition comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor
      \log_2 a \rfloor}})$} \processes. This decreases every new
partition until reaching one \process per new partition: its
source. The average number of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}, where \smash{$\overline{|O|}$} is the average
number of neighbors per \process.
% \TODO{Multiple receipt and
% multiple
%  delivery imply more messages (receipt bounded by $|O|$ as well).}
Deleting the $a^{th}$ partition generates twice as many messages as
creating the $a^{th}$ partition: $\delta$ messages travel through the
partition, then $\alpha$ messages compete to fill the gap.  Overall,
the communication complexity shows that \NAME scales well to the
number of partitions.
%% In the
%% worst-case, every new partition includes all but one \process
%% belonging to the previous partition. The total number of messages
%% after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
%% a^2)$. As for the average-case, the number of messages for the
%% partition deletion is identical to the number of messages of the
%% corresponding partition creation.
